Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-09-09 overall.tex
Iteration on the base of Sergey's comments

    \item REVIEWER 2: How does the poor metadata read performance of
    1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
    factored into the choice of walltime requests?\mtnote{Asked Sergey and
    Danila to confirm these numbers and walltime request policy.} Is it
    applicable only to TITAN, to HPC in general, or even to regular WLCG
    sites?\mtnote{This is clearly out of scope for this paper.} Are there
    solutions at application, e.g. the use of packaged python scripts and
    libraries?\mtnote{Added the following: ``While these figures are specific
    to OLCF and TItan, reengineering on AthenaMP could further improve
    startup time on every platform.''} The information presented is
    insufficient to estimate whether this issue applies to other environments
    and usecases as well, or is simply due to suboptimal
    implementation/configuration.

    \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
    ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
    operations accounts for \~10\% of the average workload (see [1], section
    ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
    figure certainly comes from averaging over many different workloads
    across all experiments, it seems a bit surprising that FP scheduling
    contention is responsible for 50\% decrease in compute efficiency on
    Titan when compared to other WLCG sites (paper page 6).\mtnote{Wrote an email to Sergey asking for a clarification.}
\end{enumerate}
\fi

@@

Initially, all the shared libraries and the python scripts of AthenaMP were
stored on the OLCF Spider 2 Lustre file system. However, the I/O patterns of
the initialization and configuration stages degraded the performance of the
filesystem. This was addressed by moving the AthenaMP distribution to a
read-only NFS directory, shared among Titan's DTNs and worker nodes. NFS
eliminated the problem of metadata contention, improving metadata read
performance from \(\approx\)6,300 seconds on Lustre to \(\approx\)1,500
seconds on NFS\@. While these figures are specific to OLCF and TItan, reengineering on AthenaMP could further improve startup time on every platform.

@@

The differences in performance between Titan and the BNL and SIGNET Grid
sites are due to the FP scheduler competition and the availability of newer
processors. The CPUs at the Grid sites have one FP scheduler per core and are
on average newer than the CPU of Titan. The heterogeneity of the Grid sites'
CPUs explain the higher performance variance we observed compared to the
performance consistency we measured on Titan. This difference in performance
is compensated by the amount of resources available on Titan (capable of
executing up to 30\%/year of the ATLAS detector simulations) and by the lack
of further resources available to WLCG\@. Also, it should be noted that Titan
is at the end of its life-cycle and that Summit, Titan's successor, will
offer cutting-edge performances.

>>>>

    \item REVIEWER 2: How does the poor metadata read performance of
    1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
    factored into the choice of walltime requests?\mtnote{Thanks to Sergey's
    input I corrected the figure previously reported.} Is it applicable only
    to TITAN, to HPC in general, or even to regular WLCG sites?\mtnote{This
    is clearly out of scope for this paper.} Are there solutions at
    application, e.g. the use of packaged python scripts and
    libraries?\mtnote{Added the following: ``While these figures are specific
    to OLCF and TItan, reengineering on AthenaMP could further improve
    startup time on every platform.''} The information presented is
    insufficient to estimate whether this issue applies to other environments
    and usecases as well, or is simply due to suboptimal
    implementation/configuration.

    \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
    ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
    operations accounts for \~10\% of the average workload (see [1], section
    ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
    figure certainly comes from averaging over many different workloads
    across all experiments, it seems a bit surprising that FP scheduling
    contention is responsible for 50\% decrease in compute efficiency on
    Titan when compared to other WLCG sites (paper page 6).\mtnote{Thanks to
    Sergey's input I downplayed the role played by FP and up-played the role
    played by new CPUs.}
\end{enumerate}
\fi

@@

Initially, all the shared libraries and the python scripts of AthenaMP were
stored on the OLCF Spider 2 Lustre file system. However, the I/O patterns of
the initialization and configuration stages degraded the performance of the
filesystem. This was addressed by moving the AthenaMP distribution to a
read-only NFS directory, shared among Titan's DTNs and worker nodes. NFS
eliminated the problem of metadata contention, improving metadata read
performance from a worse case scenario of 6,300s on Lustre to \(\approx\)225s
on NFS\@. While these figures are specific to OLCF and TItan, reengineering
of AthenaMP could further improve startup time on every platform.

@@

The differences in performance between Titan and the BNL and SIGNET Grid
sites are due to the FP scheduler competition but mainly to the availability
of newer processors. 
% The CPUs at the Grid sites have one FP scheduler per core and are on
% average newer than the CPU of Titan.
The heterogeneity of the Grid sites' CPUs explain the higher performance
variance we observed compared to the performance consistency we measured on
Titan. This difference in performance is compensated by the amount of
resources available on Titan (capable of executing up to 30\%/year of the
ATLAS detector simulations) and by the lack of further resources available to
WLCG\@. Also, it should be noted that Titan is at the end of its life-cycle
and that Summit, Titan's successor, will offer cutting-edge performances.
